---
title: "treeducken + summary stats demo"
output: html_notebook
---

R markdown notebook exploring how we could in theory explore what some of these distributions
in cophylogenetics looks like.

First off, we need to simulate some trees.

```{r}
library(treeducken)
library(dplyr)
N <- 100
# HOST PARAMETERS
host_br <- rexp(N, rate = 1)
host_dr <- host_br / 2
# SYMBIONT PARAMETERS
symb_br <- rexp(N, rate = 1)
symb_dr <- symb_br / 2
# BOTH PARAMETERS
cospec_r <- rexp(N, rate = 1)
host_exp_rate <- rexp(N, rate = 1)
cophy_set <- vector(length = N)
for(i in 1:N){
   cophy_set[[i]] <- sim_cophyBD(hbr = host_br[[i]],
                            sbr = symb_br[[i]],
                            hdr = host_dr[[i]],
                            sdr = symb_dr[[i]],
                            cosp_rate = cospec_r[[i]],
                            host_exp_rate = host_exp_rate[[i]],
                            numbsim = 1,
                            time_to_sim = 1.0)
}
```


## Tree Statistic calculation


```{r}
library(castor)
library(dplyr)
host_trees <- vector(length = length(cophy_set))
symb_trees <- vector(length = length(cophy_set))
intxn_mats <- vector(length = length(cophy_set))

# Make a list for host trees, symbiont trees, and interaction matrices
for(i in 1:N){
  host_trees[[i]] <-treeducken::host_tree(cophy_set[[i]]) 
  host_trees[[i]] <- host_trees[[i]]$host_tree %>% treeducken::drop_extinct()
  symb_trees[[i]] <- treeducken::symb_tree(cophy_set[[i]]) 
  symb_trees[[i]] <- symb_trees[[i]]$symb_tree %>% treeducken::drop_extinct()
  intxn_mats[[i]] <- treeducken::association_mat(cophy_set[[i]]) 
}

summary_stat <- data.frame()
```


Some preliminary formatting has now been done. We can actually calculate stuff now. Note that I did prune the extinct tips to better resemble empirical datasets. In the majority of cases, 

```{r}
host_tree_colless_stat <- vector(length = N)
symb_tree_colless_stat <- vector(length = N)
host_sackin_indx <- vector(length = N)
symb_sackin_indx <- vector(length = N)
host_blum_indx <- vector(length = N)
symb_blum_indx <- vector(length = N)

for(i in 1:N){
  host_colless_stat[[i]] <- host_trees[[i]] %>% tree_imbalance(type = "Colless")
  symb_colless_stat[[i]] <- symb_trees[[i]] %>% tree_imbalance(type = "Colless")
  host_sackin_indx[[i]] <- host_trees[[i]] %>% tree_imbalance(type = "Sackin")
  symb_sackin_indx[[i]] <- symb_trees[[i]] %>% tree_imbalance(type = "Sackin")
  host_blum_indx[[i]] <- host_trees[[i]] %>% tree_imbalance(type = "Blum")
  symb_blum_indx[[i]] <- symb_trees[[i]] %>% tree_imbalance(type = "Blum")

}

summary_stat <- data.frame(host_colless_stat, 
                           symb_colless_stat, 
                           host_sackin_indx, 
                           symb_sackin_indx,
                           host_blum_indx,
                           symb_blum_indx)
```

Now I have some of these tree balance statistics. Let us plot these.


```{r}
library(ggplot2)
ggplot(data = summary_stat, aes(x = host_colless_stat, y = symb_colless_stat)) + 
  geom_point() +
  ylab("Symbiont Colless") + 
  xlab("Host Colless") + 
  theme_bw()

ggplot(data = summary_stat, aes(x = host_sackin_indx, y = symb_sackin_indx)) + 
  geom_point() +
  ylab("Symbiont Colless") + 
  xlab("Host Colless") + 
  theme_bw()

ggplot(data = summary_stat, aes(x = host_blum_indx, y = symb_blum_indx)) + 
  geom_point() +
  ylab("Symbiont Colless") + 
  xlab("Host Colless") + 
  theme_bw()
```